R Notebook

### 2021-03-12 examples/Toy-example-balanced-accessibility.Rmd
Update accessibility calculations

dist_threshold <- quantile(as.vector(ttm_walk$travel_time), probs = c(0, 1/10, 3/10, 5/10, 7/10, 8/10, 9/10))[[5]]
```

@@

@@

Calculate the sum of the impedance by SoBi hubs for the balancing factors. In the binary case, this should be the same as the number of interpolated population unit that each SoBi hub serves; when the impedance is not binary, this is the sum of all impedance values:
```{r}
sum_b2 <- ttm_walk %>%
  group_by(hub) %>%
  summarize(sum_b2 = sum(impedance_binary))
sum_b2
```

@@

Same thing for the other accessibility estimates. Spatial join:
```{r}
accessibility_da <- population_sobi %>%
  st_join(hamilton_da_sobi %>% dplyr::select(DA))
```

@@

Retrieve data:
```{r eval=FALSE}
mhi <- get_census(dataset='CA16', 
                          regions=list(CMA=c("35537")),
                          vectors=c("v_CA16_2397"),
                          level='DA', 
                          use_cache = FALSE)
```

@@

Plot:
```{r}
 map <- ggplot() +
  geom_sf(data = data, mapping = aes(fill = bi_class), color = "white", size = 0.1, show.legend = FALSE) +
  bi_scale_fill(pal = "DkBlue", dim = 3) +
  bi_theme()

@@

Finalize plot:
```{r}
 cowplot::ggdraw() +
  cowplot::draw_plot(map, 0, 0, 1, 1) +
  cowplot::draw_plot(legend, 0.05, .6, 0.2, 0.2)
```

>>>>

# dist_threshold <- quantile(as.vector(ttm_walk$travel_time), probs = c(0, 1/10, 3/10, 5/10, 7/10, 8/10, 9/10))[[5]]

# distance threshold has to be less than 30, maximum distance people are willing to go is distance threshold

dist_threshold <- 10
```

@@

Our calculations assume that 70% of population cells have at least one station within 24 min to reach a SoBi station. We set the threshold (what people are willing to travel). Gaussian says people are more willing to go 5 than 10, less willing to go 15 than 20. Progressively discount, but we don't know the willingness. If we had more information about how many people go 5, 10, etc. then using the Gaussian function would be possible. Otherwise, there are a lot of assumptions. Binary requires us to make fewer assumptions.

@@

Calculate the sum of the impedance by SoBi hubs for the balancing factors. In the binary case, this should be the same as the number of interpolated population cells that each SoBi hub serves; when the impedance is not binary, this is the sum of all impedance values:
```{r}
sum_b2 <- ttm_walk %>%
  group_by(hub) %>%
  summarize(sum_b2 = sum(impedance_binary))
sum_b2
```

Hub 1 serves 1502 population cells, hub 2 serves 1393 population cells, etc.

@@

Same thing for the other accessibility estimates for the system with ERI hubs. Spatial join:
```{r}
accessibility_da <- population_sobi %>%
  st_join(hamilton_da_sobi %>% dplyr::select(DA))
```

@@

Retrieve data:
```{r eval}
mhi <- get_census(dataset='CA16', 
                          regions=list(CMA=c("35537")),
                          vectors=c("v_CA16_2397"),
                          level='DA', 
                          use_cache = FALSE)
```

@@

Plot accessibility of original system without ERI hubs:
```{r}
 map <- ggplot() +
  geom_sf(data = data, mapping = aes(fill = bi_class), color = "white", size = 0.1, show.legend = FALSE) +
  bi_scale_fill(pal = "DkBlue", dim = 3) +
  bi_theme()

@@

Finalize plot of accessibility with ERI hubs:
```{r}
 cowplot::ggdraw() +
  cowplot::draw_plot(map, 0, 0, 1, 1) +
  cowplot::draw_plot(legend, 0.05, .6, 0.2, 0.2)
```

Create a class for bivariate plotting (drop DA with missing income data):
```{r}
data_equity <- bi_class(accessibility_da %>%
                   drop_na(),
                 x = accessibility, 
                 y = median_household_income, 
                 style = "quantile", 
                 dim = 3) %>%
  st_as_sf()
```

Plot accessibility of original system with ERI hubs:
```{r}
 map_equity <- ggplot() +
  geom_sf(data = data_equity, mapping = aes(fill = bi_class), color = "white", size = 0.1, show.legend = FALSE) +
  bi_scale_fill(pal = "DkBlue", dim = 3) +
  bi_theme()

map_equity
```

Create legend:
```{r}
legend_equity <- bi_legend(pal = "DkBlue",
                    dim = 3,
                    xlab = "Higher accessibility ",
                    ylab = "Higher income",
                    size = 6)
```

Finalize plot of accessibility with ERI hubs:
```{r}
 cowplot::ggdraw() +
  cowplot::draw_plot(map_equity, 0, 0, 1, 1) +
  cowplot::draw_plot(legend, 0.05, .6, 0.2, 0.2)
```

Save data files:
```{r}
save(accessibility_da, file = "accessibility.RData")
save(map, file = "accessibility_map_original.RData")
save(map_equity, file = "accessibility_map_equity.RData")
save(mhi, file = "mhi.RData")
save(population_sobi_original, file = "population_sobi_original.RData")
```
